---
title: "Follow_up_analysis"
output: html_document
date: "2025-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(readr)
library(lmtest)
library(car)
library(bbmle)
library(googlesheets4)
```

#### Load data from google sheets

```{r}
url <- "https://docs.google.com/spreadsheets/d/1ljVYx-veRn9V6ruSoH4KI46yF5Y4VarCVwQGWmdjz8w/edit?gid=673524059#gid=673524059"
week1 <- read_sheet(url, sheet="Germ_Week1")
week2 <- read_sheet(url, sheet="Germ_Week2")

```

#### Combine weekly data into 1 tibble

```{r}
week1 <- week1[,1:14]
week2 <- week2 %>%
  select(Pop, Treatment, pos_tray, 8:14)
         
two_weeks <- week1 %>%
  left_join(week2, by=c("Pop", "Treatment", "pos_tray"))
  
```

#### Check germination count does not exceed actual count

```{r}
count_sum <- two_weeks %>%
  mutate(count = rowSums(select(.,8:ncol(two_weeks)), na.rm=TRUE),
         exceeds_count = count > Actual_count)
```

#### Pivot longer, convert date to days in imbibing conditions

```{r}
two_weeks_long <- two_weeks %>%
  pivot_longer(cols= 8:ncol(two_weeks),
               names_to = "Date",
               values_to = "Num_germ") %>%
  mutate(Date = mdy(Date))

start_date <- as.Date("2025-02-12")

two_weeks_long <- two_weeks_long %>%
  mutate(Days_to_Germ = as.integer(Date-start_date))
```

#### Add cumulative sum per cell

```{r}
two_weeks_long <- two_weeks_long %>%
  group_by(pos_tray) %>%
  arrange(Days_to_Germ) %>%
  mutate(Cum_germ = cumsum(Num_germ)) %>%
  ungroup()
```

#### Calculate germination proportion per cell by day

```{r}
two_weeks_long <- two_weeks_long %>%
  mutate(Germ_prop = Cum_germ/Actual_count)
```

#### Calculate total germination rate per cell

```{r}
two_weeks_long <- two_weeks_long %>%
  group_by(pos_tray) %>%
  arrange(Days_to_Germ) %>% 
  mutate(Germ_rate = sum(Num_germ) / sum(Num_germ * Days_to_Germ),
       Germ_rate = ifelse(is.nan(Germ_rate) | is.infinite(Germ_rate), 
                                 0, Germ_rate)) %>%
  ungroup()
```

#### Calculating averages for each pop and treatment

```{r}
two_weeks_avg <- two_weeks_long %>%
  group_by(Pop, Treatment, Days_to_Germ) %>%
  summarize(Avg_germ_prop = mean(Germ_prop, na.rm=TRUE),
            SD_germ_prop = sd(Germ_prop, na.rm=TRUE),
            Avg_germ_rate = mean(Germ_rate, na.rm=TRUE),
            SD_germ_rate = sd(Germ_rate, na.rm=TRUE),
            .groups="drop")
```

#### Add elevation data

```{r}
elevation <- read_sheet(url, sheet="Elevation")

two_weeks_long <- two_weeks_long %>%
  left_join(elevation, by="Pop")

two_weeks_avg <- two_weeks_avg %>%
  left_join(elevation, by="Pop")

#add levels to pop so we can order by elevation
two_weeks_avg <- two_weeks_avg %>%
  arrange(Elev) %>%
  mutate(Pop = factor(Pop, levels = unique(Pop)))

two_weeks_long <- two_weeks_long %>%
  arrange(Elev) %>%
  mutate(Pop = factor(Pop, levels=unique(Pop)))
```

#### Average germination prop by elevation

```{r}
two_weeks_avg %>%
  filter(Days_to_Germ == 14) %>%
  ggplot(aes(x=Elev, y=Avg_germ_prop, color=Treatment)) +
  geom_point()+
  geom_line()
```

#### Germ prop ordered by elevation

```{r}
ggplot(two_weeks_avg, aes(x=Days_to_Germ)) +
  geom_line(aes(y=Avg_germ_prop, color=Treatment), linewidth=1) +
  geom_ribbon(aes(ymin = Avg_germ_prop - SD_germ_prop, 
                  ymax = Avg_germ_prop + SD_germ_prop, 
                  fill = Treatment), 
              alpha = 0.1) +
  theme_bw()+
  facet_wrap(.~Pop, ncol=4, scales="free")+
  ylab("Average Germination Proportion") +
  xlab("Days in Inductive Conditions")
```

#### Plot germ rate by pop and treatment ordered by elevation

```{r}
ggplot(two_weeks_long, aes(x=Treatment, y=Germ_rate, color=Treatment)) +
  geom_boxplot()+
  theme_bw()+
  facet_wrap(.~Pop, ncol=4, scales="free_y")+
  ylab("Average Germination Rate") 
```

#### GLMs

```{r}
end_two_weeks <- two_weeks_long %>%
  filter(Days_to_Germ == 14)

m1 <- glm(cbind(Cum_germ, Actual_count) ~ 0 + Pop * Treatment,
                        data = end_two_weeks, 
                        family = binomial())
m1null <- glm(cbind(Cum_germ, Actual_count) ~ 1,
                        data = end_two_weeks, 
                        family = binomial())
summary(m1)
lrtest(m1null, m1)
Anova(m1) #population alone is the only significant fixed effect, this is different from the week 1 data there the interaction was also significant

m2 <- glm(cbind(Cum_germ, Actual_count) ~ 0 + Elev * Treatment,
                        data = end_two_weeks, 
                        family = binomial())
summary(m2)
lrtest(m1null, m2)
Anova(m2)
#anova(m2, test="Chisq") seems to show the same significance as marginal hypothesis test

ICtab(m1, m2) #population is a better fit than elevation...
```
