---
title: "Follow_up_analysis"
output: html_document
date: "2025-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(readr)
library(lmtest)
library(car)
library(bbmle)
library(googlesheets4)
library(lme4)
```

#### Load data from google sheets

```{r}
url <- "https://docs.google.com/spreadsheets/d/1ljVYx-veRn9V6ruSoH4KI46yF5Y4VarCVwQGWmdjz8w/edit?gid=673524059#gid=673524059"
week1 <- read_sheet(url, sheet="Germ_Week1")
week2 <- read_sheet(url, sheet="Germ_Week2")
week3 <- read_sheet(url, sheet="Germ_Week3")
week4 <- read_sheet(url, sheet="Germ_Week4")

```

#### Combine weekly data into 1 tibble

```{r}
week1 <- week1[,1:14]
week2 <- week2 %>%
  select(Pop, Treatment, pos_tray, 8:14)
week3 <- week3 %>%
  select(Pop, Treatment, pos_tray, 8:10)
week4 <- week4 %>%
  select(Pop, Treatment, pos_tray, 8:11)
         
two_weeks <- week1 %>%
  left_join(week2, by=c("Pop", "Treatment", "pos_tray"))

four_weeks <- two_weeks %>%
  left_join(week3, by=c("Pop", "Treatment", "pos_tray")) %>%
  left_join(week4, by=c("Pop", "Treatment", "pos_tray"))
  
```

#### Check germination count does not exceed actual count

```{r}
count_sum <- two_weeks %>%
  mutate(count = rowSums(select(.,8:ncol(two_weeks)), na.rm=TRUE),
         exceeds_count = count > Actual_count)

count_sum4 <- four_weeks %>%
  mutate(count = rowSums(select(.,8:ncol(four_weeks)), na.rm=TRUE),
         exceeds_count = count > Actual_count)
```

#### Pivot longer, convert date to days in imbibing conditions

```{r}
two_weeks_long <- two_weeks %>%
  pivot_longer(cols= 8:ncol(two_weeks),
               names_to = "Date",
               values_to = "Num_germ") %>%
  mutate(Date = mdy(Date))

start_date <- as.Date("2025-02-12")

two_weeks_long <- two_weeks_long %>%
  mutate(Days_to_Germ = as.integer(Date-start_date))

four_weeks_long <- four_weeks %>%
  pivot_longer(cols= 8:ncol(four_weeks),
               names_to = "Date",
               values_to = "Num_germ") %>%
  mutate(Date = mdy(Date))

four_weeks_long <- four_weeks_long %>%
  mutate(Days_to_Germ = as.integer(Date-start_date))
```

#### Add cumulative sum per cell

```{r}
two_weeks_long <- two_weeks_long %>%
  group_by(pos_tray) %>%
  arrange(Days_to_Germ) %>%
  mutate(Cum_germ = cumsum(Num_germ)) %>%
  ungroup()

four_weeks_long <- four_weeks_long %>%
  group_by(pos_tray) %>%
  arrange(Days_to_Germ) %>%
  mutate(Cum_germ = cumsum(Num_germ)) %>%
  ungroup()
```

#### Calculate germination proportion per cell by day

```{r}
two_weeks_long <- two_weeks_long %>%
  mutate(Germ_prop = Cum_germ/Actual_count)

four_weeks_long <- four_weeks_long %>%
  mutate(Germ_prop = Cum_germ/Actual_count)
```

#### Calculate germination rate per seed

```{r}
#need to create new rows for each seed that germinated since multiple seeds germinated in the same cell on the same day
two_weeks_no_germ <- two_weeks_long %>%
  filter(Num_germ == 0) %>%
  mutate(Germ_rate_seed = NA_real_)

two_weeks_germ <- two_weeks_long %>%
  filter(Num_germ > 0) %>%
  mutate(Num_germ_original = Num_germ) %>%
  uncount(Num_germ) %>%  # Expand the rows based on the number of seeds germinated
  mutate(Germ_rate_seed = 1 / Days_to_Germ)

two_weeks_germ <- two_weeks_germ %>%
  mutate(Num_germ = Num_germ_original)

two_weeks_long_expand <- bind_rows(two_weeks_no_germ, two_weeks_germ)
two_weeks_long_expand <- two_weeks_long_expand[,1:13]



four_weeks_no_germ <- four_weeks_long %>%
  filter(Num_germ == 0) %>%
  mutate(Germ_rate_seed = NA_real_)

four_weeks_germ <- four_weeks_long %>%
  filter(Num_germ > 0) %>%
  mutate(Num_germ_original = Num_germ) %>%
  uncount(Num_germ) %>%  # Expand the rows based on the number of seeds germinated
  mutate(Germ_rate_seed = 1 / Days_to_Germ)

four_weeks_germ <- four_weeks_germ %>%
  mutate(Num_germ = Num_germ_original)

four_weeks_long_expand <- bind_rows(four_weeks_no_germ, four_weeks_germ)
four_weeks_long_expand <- four_weeks_long_expand[,1:13]


#two_weeks_long <- two_weeks_long %>%
 # group_by(pos_tray) %>%
 # arrange(Days_to_Germ) %>% 
 # mutate(Germ_rate = sum(Num_germ) / sum(Num_germ * Days_to_Germ),
     #  Germ_rate = ifelse(is.nan(Germ_rate) | is.infinite(Germ_rate), 
       #                          0, Germ_rate)) %>%
  #ungroup()

#four_weeks_long <- four_weeks_long %>%
 # group_by(pos_tray) %>%
 # arrange(Days_to_Germ) %>% 
 # mutate(Germ_rate = sum(Num_germ) / sum(Num_germ * Days_to_Germ),
  #     Germ_rate = ifelse(is.nan(Germ_rate) | is.infinite(Germ_rate), 
  #                               0, Germ_rate)) %>%
  #ungroup()
```

#### Calculate average germination prop and rate per day

```{r}
two_weeks_avg_prop <- two_weeks_long %>%
  group_by(Pop, Treatment, Days_to_Germ) %>%
  summarize(Avg_germ_prop = mean(Germ_prop, na.rm=TRUE),
            SD_germ_prop = sd(Germ_prop, na.rm=TRUE),
            n=sum(Germ_prop != 0),
            SE_germ_prop=SD_germ_prop/sqrt(n),
            .groups="drop")

four_weeks_avg_prop <- four_weeks_long %>%
  group_by(Pop, Treatment, Days_to_Germ) %>%
  summarize(Avg_germ_prop = mean(Germ_prop, na.rm=TRUE),
            SD_germ_prop = sd(Germ_prop, na.rm=TRUE),
            n=sum(Germ_prop != 0),
            SE_germ_prop=SD_germ_prop/sqrt(n),
            .groups="drop")


two_weeks_avg_rate <- two_weeks_long_expand %>%
  group_by(Pop, Treatment) %>%
  summarize(Avg_germ_rate = mean(Germ_rate_seed, na.rm=TRUE),
            SD_germ_rate = sd(Germ_rate_seed, na.rm=TRUE),
            n=sum(!is.na(Germ_rate_seed)),
            SE_germ_rate=SD_germ_rate/sqrt(n),
            .groups="drop")

four_weeks_avg_rate <- four_weeks_long_expand %>%
  group_by(Pop, Treatment) %>%
  summarize( Avg_germ_rate = mean(Germ_rate_seed, na.rm=TRUE),
            SD_germ_rate = sd(Germ_rate_seed, na.rm=TRUE),
            n=sum(!is.na(Germ_rate_seed)),
            SE_germ_rate=SD_germ_rate/sqrt(n),
            .groups="drop")
```

#### Add elevation and latitude data - need to run each line individually for some reason

```{r}
elevation <- read_sheet(url, sheet="Elevation_Latitude")

two_weeks_long_expand <- two_weeks_long_expand %>%
  left_join(elevation, by="Pop")

two_weeks_long <- two_weeks_long%>%
  left_join(elevation, by="Pop")

two_weeks_avg_prop <- two_weeks_avg_prop %>%
  left_join(elevation, by="Pop")

two_weeks_avg_rate <- two_weeks_avg_rate %>%
  left_join(elevation, by="Pop")


four_weeks_long_expand <- four_weeks_long_expand %>%
  left_join(elevation, by="Pop")

four_weeks_avg_prop <- four_weeks_avg_prop %>%
  left_join(elevation, by="Pop")

four_weeks_avg_rate <- four_weeks_avg_rate %>%
  left_join(elevation, by="Pop")

#add levels to pop so we can order by elevation
two_weeks_avg_prop <- two_weeks_avg_prop %>%
  arrange(Elev) %>%
  mutate(Pop = factor(Pop, levels = unique(Pop)))

two_weeks_avg_rate <- two_weeks_avg_rate %>%
  arrange(Elev) %>%
  mutate(Pop = factor(Pop, levels = unique(Pop)))

two_weeks_long_expand <- two_weeks_long_expand %>%
  arrange(Elev) %>%
  mutate(Pop = factor(Pop, levels=unique(Pop)))

#two_weeks_avg <- two_weeks_avg %>%
 # arrange(Lat) %>%
 # mutate(Pop = factor(Pop, levels = unique(Pop)))

#two_weeks_long <- two_weeks_long %>%
 # arrange(Lat) %>%
 # mutate(Pop = factor(Pop, levels=unique(Pop)))

four_weeks_avg_prop <- four_weeks_avg_prop %>%
  arrange(Elev) %>%
  mutate(Pop = factor(Pop, levels = unique(Pop)))

four_weeks_avg_rate <- four_weeks_avg_rate %>%
  arrange(Elev) %>%
  mutate(Pop = factor(Pop, levels = unique(Pop)))

four_weeks_long_expand <- four_weeks_long_expand %>%
  arrange(Elev) %>%
  mutate(Pop = factor(Pop, levels=unique(Pop)))
```

#### Average germination prop by elevation

```{r}
library(pscl)   
end_week2 <- two_weeks_long_expand %>%
  filter(Days_to_Germ == 14)

#test if germ_prop differed for indiv seeds by elev*treatment with cell nested within rep, need to use a 0-inflated distribution

#end_week2$predicted <- predict(m1, type = "response", newdata = final_count)

two_weeks_avg_prop %>%
  filter(Days_to_Germ == 14) %>%
  ggplot(aes(x=Elev, y=Avg_germ_prop, color=Treatment)) +
  geom_point() 
  #geom_line(data = end_week2, aes(x = Elev#, y = predicted, color = Treatment), 
  #          linetype = "solid")

#two_weeks_avg %>%
 # filter(Days_to_Germ == 14) %>%
 # ggplot(aes(x=Lat, y=Avg_germ_prop, color=Treatment)) +
  #geom_point()+
 # geom_line()

four_weeks_avg_prop %>%
  filter(Days_to_Germ == 33) %>%
  ggplot(aes(x=Elev, y=Avg_germ_prop, color=Treatment)) +
 geom_point()

#four_weeks_avg %>%
  #filter(Days_to_Germ == 14) %>%
  #ggplot(aes(x=Lat, y=Avg_germ_prop, color=Treatment)) +
  #geom_point()+
  #geom_line()

```

So if you are modeling germination rate, you could either model it for each seed and have well as a random effect which would be nested within block (I think) or you could calculate average germination rate for each well, based on individual seed germination, and use well-level germination rate as the response variable and just have block as the random effect.

#### Germ prop ordered by elevation - TRY TO ORDER BY LATITUDE

```{r}
ggplot(two_weeks_avg_prop, aes(x=Days_to_Germ, y=Avg_germ_prop, color=Treatment)) +
  geom_point() +
  geom_line()+
 # geom_line(aes(y=Avg_germ_prop, color=Treatment), linewidth=1) +
 # geom_ribbon(aes(ymin = Avg_germ_prop - SE_germ_prop, 
  #                ymax = Avg_germ_prop + SE_germ_prop, 
  #                fill = Treatment), 
  #            alpha = 0.2) +
  theme_bw()+
  facet_wrap(.~Pop, ncol=4, scales="free")+
  ylab("Average Germination Proportion") +
  xlab("Days in Inductive Conditions")

ggplot(four_weeks_avg_prop, aes(x=Days_to_Germ, y=Avg_germ_prop, color=Treatment)) +
  geom_point() +
  geom_line()+
 # geom_line(aes(y=Avg_germ_prop, color=Treatment), linewidth=1) +
  #geom_ribbon(aes(ymin = Avg_germ_prop - SE_germ_prop, 
   #               ymax = Avg_germ_prop + SE_germ_prop, 
    #              fill = Treatment), 
     #         alpha = 0.1) +
  theme_bw()+
  facet_wrap(.~Pop, ncol=4, scales="free")+
  ylab("Average Germination Proportion") +
  xlab("Days in Inductive Conditions")
```

#### Plot germ rate by pop and treatment ordered by elevation

```{r}
ggplot(two_weeks_avg_rate, aes(x=Treatment, y=Avg_germ_rate, color=Treatment)) +
  geom_point()+
  geom_errorbar(aes(ymin = Avg_germ_rate-SE_germ_rate, ymax=Avg_germ_rate+SE_germ_rate), width=0.2) +
  theme_bw()+
  facet_wrap(.~Pop, ncol=4, scales="free_y")+
  theme(axis.text.x=element_text(angle=90, hjust=1)) +
  ylab("Average Germination Rate") 

ggplot(four_weeks_long, aes(x=Treatment, y=Germ_rate, color=Treatment)) +
  geom_boxplot()+
  theme_bw()+
  facet_wrap(.~Pop, ncol=4, scales="free_y")+
  ylab("Average Germination Rate") 
```

#### GLMs

```{r}
end_two_weeks <- two_weeks_long %>%
  filter(Days_to_Germ == 14)

#
m1 <- glm(cbind(Cum_germ, Actual_count-Cum_germ) ~ 0 + Pop * Treatment,
                        data = end_two_weeks, 
                        family = binomial())
m1null <- glm(cbind(Cum_germ, Actual_count-Cum_germ) ~ 1,
                        data = end_two_weeks, 
                        family = binomial())
summary(m1)
lrtest(m1null, m1)
Anova(m1) #population alone is the only significant fixed effect, this is different from the week 1 data there the interaction was also significant

m2 <- glm(cbind(Cum_germ, Actual_count) ~ 0 + Elev * Treatment,
                        data = end_two_weeks, 
                        family = binomial())
summary(m2)
lrtest(m1null, m2)
Anova(m2)
#anova(m2, test="Chisq") seems to show the same significance as marginal hypothesis test

ICtab(m1, m2) #population is a better fit than elevation...


#mixed models
m3 <-  glmer(cbind(Cum_germ, Actual_count-Cum_germ) ~ -1 +Treatment + (1 | Pop), 
               data = end_two_weeks, 
               family = binomial)
Anova(m3)
summary(m3)
```
