---
title: "Maya Analysis"
output: html_document
date: "2025-01-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(readr)
```

Import data
```{r}
wk1 <- read_csv("../Input/Week1_Germ.csv")
wk2 <- read_csv("../Input/week2_germ.csv")
wk3 <- read_csv("../Input/Week3_Germ.csv")
wk4 <- read_csv("../Input/week4_germ.csv")
```
Combine all 4 weeks into one tibble
```{r}
wk2_date <- wk2 %>%
  select(Pop, Treatment, Cell, 7:ncol(wk2))
wk3_date <- wk3 %>%
  select(Pop, Treatment, Cell, 8:ncol(wk3))
wk4_date <- wk4 %>%
  select(Pop, Treatment, Cell, 8:ncol(wk4))

all_wk <- wk1 %>%
  left_join(wk2_date, by=c("Pop", "Treatment", "Cell")) %>%
  left_join(wk3_date, by=c("Pop", "Treatment", "Cell")) %>%
  left_join(wk4_date, by=c("Pop", "Treatment", "Cell"))
```

check actual count matches across weeks
```{r}
wk1_act_count <- wk1 %>%
  select(Pop, Treatment, Cell, Actual_count)
wk3_act_count <- wk3 %>%
  select(Pop, Treatment, Cell, Actual_count)
wk4_act_count <- wk4 %>%
  select(Pop, Treatment, Cell, Actual_count)

actual_count <- wk1_act_count %>%
  left_join(wk3_act_count, by=c("Pop", "Treatment", "Cell")) %>%
  left_join(wk4_act_count, by=c("Pop", "Treatment", "Cell")) %>%
  mutate(check_count = ifelse(Actual_count.x == Actual_count.y & Actual_count.y == Actual_count, TRUE, FALSE))

count_sum <- all_wk %>%
  mutate(count = rowSums(select(.,8:ncol(all_wk)), na.rm=TRUE),
         exceeds_count = count > Actual_count)
```

Two rows have total counts that exceed the actual count - update the actual count to match the total 
```{r}
all_wk <- all_wk %>%
  mutate(Actual_count = case_when(
    Tray == 8 & Rep == 4 & Treatment == "Control" & Pop == "IH" & Cell == "A4_T8" ~ 6,
    Tray == 6 & Rep == 3 & Treatment == "05KNO3" & Pop == "LVTR" & Cell == "A6_T6" ~ 6,
      TRUE ~ Actual_count 
    )
  )
```

Pivot longer, convert date to days in imbibing conditions
```{r}
all_wk_long <- all_wk %>%
  pivot_longer(cols= 8:ncol(all_wk),
               names_to = "Date",
               values_to = "Num_germ") %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

start_date <- as.Date("2024-07-23")

all_wk_long <- all_wk_long %>%
  mutate(Days_to_Germ = as.integer(Date-start_date))
```

Add cumulative sum 
```{r}
all_wk_long <- all_wk_long %>%
  group_by(Cell) %>%
  arrange(Days_to_Germ) %>%
  mutate(Cum_germ = cumsum(Num_germ)) %>%
  ungroup()
```

Calculate germination proportion per cell by day
```{r}
all_wk_long <- all_wk_long %>%
  mutate(Germ_prop = Cum_germ/Expect_count)
```

Calculate total germination rate per cell
```{r}
all_wk_long <- all_wk_long %>%
  group_by(Cell) %>%
  arrange(Days_to_Germ) %>% 
  mutate(Germ_rate = sum(Num_germ) / sum(Num_germ * Days_to_Germ),
       Germ_rate = ifelse(is.nan(Germ_rate) | is.infinite(Germ_rate), 
                                 0, Germ_rate)) %>%
  ungroup()
```

Calculate daily germination rate per cell
```{r}
all_wk_long <- all_wk_long %>%
  group_by(Cell) %>%
  arrange(Days_to_Germ) %>% 
mutate(
  Germ_rate2 = Num_germ / Actual_count,
  Germ_rate2 = ifelse(is.nan(Germ_rate2) | is.infinite(Germ_rate2), 0, Germ_rate2)
  ) %>%
  ungroup()
```

Calculating averages for each pop and treatment
```{r}
all_wk_avg <- all_wk_long %>%
  group_by(Pop, Treatment, Days_to_Germ) %>%
  summarize(Avg_germ_prop = mean(Germ_prop, na.rm=TRUE),
            SD_germ_prop = sd(Germ_prop, na.rm=TRUE),
            Avg_germ_rate = mean(Germ_rate, na.rm=TRUE),
            SD_germ_rate = sd(Germ_rate, na.rm=TRUE),
            Avg_germ_rate2 = mean(Germ_rate2, na.rm=TRUE),
            SD_germ_rate2 = sd(Germ_rate2, na.rm=TRUE),
            .groups="drop")

```

Plot averages by pop and treatment
```{r}
ggplot(all_wk_avg, aes(x=Days_to_Germ)) +
  geom_line(aes(y=Avg_germ_prop, color=Treatment), size=1) +
  geom_ribbon(aes(ymin = Avg_germ_prop - SD_germ_prop, 
                  ymax = Avg_germ_prop + SD_germ_prop, 
                  fill = Treatment), 
              alpha = 0.1) +
  theme_bw()+
  facet_wrap(.~Pop, ncol=4)+
  ylab("Average Germination Proportion") +
  xlab("Days in Inductive Conditions")
```
```{r}
ggplot(all_wk_long, aes(x=Treatment, y=Germ_rate, color=Treatment)) +
  geom_boxplot()+
  theme_bw()+
  facet_wrap(.~Pop, ncol=4, scales="free_y")+
  ylab("Average Germination Rate") 
```
```{r}
ggplot(all_wk_avg, aes(x=Days_to_Germ)) +
  geom_line(aes(y=Avg_germ_rate2, color=Treatment), size=0.7) +
  geom_ribbon(aes(ymin = Avg_germ_rate2 - SD_germ_rate2, 
                  ymax = Avg_germ_rate2 + SD_germ_rate2, 
                  fill = Treatment), 
              alpha = 0.1) +
  theme_bw()+
  facet_wrap(.~Pop, ncol=4, scales= "free_y")+
  scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
  ylab("Average Germination Rate2") +
  xlab("Days in Inductive Conditions")
```
Add elevation data
```{r}
elev <- read_csv("../input/elevation_data.csv")

all_wk_long <- all_wk_long %>%
  left_join(elev, by="Pop")

all_wk_avg <- all_wk_avg %>%
  left_join(elev, by="Pop")


```
Average germination prop by elevation
```{r}
all_wk_avg %>%
  filter(Days_to_Germ == 42) %>%
  ggplot(aes(x=Elev, y=Avg_germ_prop, color=Treatment)) +
  geom_point()+
  geom_line()
```

