---
title: "Maya Analysis"
output: html_document
date: "2025-01-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(readr)
library(lmtest)
library(car)
library(bbmle)
```

#### Import Data from all weeks

```{r}
wk1 <- read_csv("../Input/Week1_Germ.csv")
wk2 <- read_csv("../Input/week2_germ.csv")
wk3 <- read_csv("../Input/Week3_Germ.csv")
wk4 <- read_csv("../Input/week4_germ.csv") #moreso week 4+
```

#### Combine all 4 weeks into one tibble

```{r}
wk2_date <- wk2 %>%
  select(Pop, Treatment, Cell, 7:ncol(wk2))
wk3_date <- wk3 %>%
  select(Pop, Treatment, Cell, 8:ncol(wk3))
wk4_date <- wk4 %>%
  select(Pop, Treatment, Cell, 8:ncol(wk4))

all_wk <- wk1 %>%
  left_join(wk2_date, by=c("Pop", "Treatment", "Cell")) %>%
  left_join(wk3_date, by=c("Pop", "Treatment", "Cell")) %>%
  left_join(wk4_date, by=c("Pop", "Treatment", "Cell"))
```

#### Check actual count matches across weeks

```{r}
wk1_act_count <- wk1 %>%
  select(Pop, Treatment, Cell, Actual_count)
wk3_act_count <- wk3 %>%
  select(Pop, Treatment, Cell, Actual_count)
wk4_act_count <- wk4 %>%
  select(Pop, Treatment, Cell, Actual_count)

actual_count <- wk1_act_count %>%
  left_join(wk3_act_count, by=c("Pop", "Treatment", "Cell")) %>%
  left_join(wk4_act_count, by=c("Pop", "Treatment", "Cell")) %>%
  mutate(check_count = ifelse(Actual_count.x == Actual_count.y & Actual_count.y == Actual_count, TRUE, FALSE))

count_sum <- all_wk %>%
  mutate(count = rowSums(select(.,8:ncol(all_wk)), na.rm=TRUE),
         exceeds_count = count > Actual_count)
```

##### Two rows have total counts that exceed the actual count - update the actual count to match the total

```{r}
all_wk <- all_wk %>%
  mutate(Actual_count = case_when(
    Tray == 8 & Rep == 4 & Treatment == "Control" & Pop == "IH" & Cell == "A4_T8" ~ 6,
    Tray == 6 & Rep == 3 & Treatment == "05KNO3" & Pop == "LVTR" & Cell == "A6_T6" ~ 6,
      TRUE ~ Actual_count 
    )
  )
```

#### Pivot longer, convert date to days in imbibing conditions

```{r}
all_wk_long <- all_wk %>%
  pivot_longer(cols= 8:ncol(all_wk),
               names_to = "Date",
               values_to = "Num_germ") %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

start_date <- as.Date("2024-07-23")

all_wk_long <- all_wk_long %>%
  mutate(Days_to_Germ = as.integer(Date-start_date))
```

#### Add cumulative sum per cell

```{r}
all_wk_long <- all_wk_long %>%
  group_by(Cell) %>%
  arrange(Days_to_Germ) %>%
  mutate(Cum_germ = cumsum(Num_germ)) %>%
  ungroup()
```

#### Calculate germination proportion per cell by day

```{r}
all_wk_long <- all_wk_long %>%
  mutate(Germ_prop = Cum_germ/Actual_count)
```

#### Calculate total germination rate per cell

```{r}
all_wk_long <- all_wk_long %>%
  group_by(Cell) %>%
  arrange(Days_to_Germ) %>% 
  mutate(Germ_rate = sum(Num_germ) / sum(Num_germ * Days_to_Germ),
       Germ_rate = ifelse(is.nan(Germ_rate) | is.infinite(Germ_rate), 
                                 0, Germ_rate)) %>%
  ungroup()
```

##### Calculate daily germination rate per cell - DON'T NEED

```{r}
#all_wk_long <- all_wk_long %>%
  #group_by(Cell) %>%
  #arrange(Days_to_Germ) %>% 
#mutate(
  #Germ_rate2 = Num_germ / Actual_count,
  #Germ_rate2 = ifelse(is.nan(Germ_rate2) | is.infinite(Germ_rate2), 0, Germ_rate2)
  #) %>%
 # ungroup()
```

#### Calculating averages for each pop and treatment

```{r}
all_wk_avg <- all_wk_long %>%
  group_by(Pop, Treatment, Days_to_Germ) %>%
  summarize(Avg_germ_prop = mean(Germ_prop, na.rm=TRUE),
            SD_germ_prop = sd(Germ_prop, na.rm=TRUE),
            Avg_germ_rate = mean(Germ_rate, na.rm=TRUE),
            SD_germ_rate = sd(Germ_rate, na.rm=TRUE),
            .groups="drop")

```

#### Add total germination count per Pop-Treatment

```{r}
#This is the same as what I calcualted above since it calculates the average of the cumulative sum for each pop-treatment combo by day
#Cum_germ is the cumulative germination count per cell and grouping by pop and treatment gets all of the cells with the same pop and treatment, and grouping by date calculates a total average for each date
all_wk_long <- all_wk_long %>%
  group_by(Date, Pop, Treatment) %>%
  mutate(Avg_total = mean(Cum_germ)) %>%
  ungroup()
```

#### Add elevation data

```{r}
elev <- read_csv("../input/elevation_data.csv")

all_wk_long <- all_wk_long %>%
  left_join(elev, by="Pop")

all_wk_avg <- all_wk_avg %>%
  left_join(elev, by="Pop")

#add levels to pop so we can order by elevation
all_wk_avg <- all_wk_avg %>%
  arrange(Elev) %>%
  mutate(Pop = factor(Pop, levels = unique(Pop)))

all_wk_long <- all_wk_long %>%
  arrange(Elev) %>%
  mutate(Pop = factor(Pop, levels=unique(Pop)))
```

#### Grouped Regression

```{r}
final_count <- all_wk_long %>%
  filter(Days_to_Germ == 42)

m1 <- glm(cbind(Cum_germ, Actual_count) ~ 0 + Pop * Treatment,
                        data = final_count, 
                        family = binomial())
m1null <- glm(cbind(Cum_germ, Actual_count) ~ 1,
                        data = final_count, 
                        family = binomial())
summary(m1)
lrtest(m1null, m1)
Anova(m1)

m2 <- glm(cbind(Cum_germ, Actual_count) ~ 0 + Elev * Treatment,
                        data = final_count, 
                        family = binomial())
summary(m2)
lrtest(m1null, m2)
Anova(m2)
#anova(m2, test="Chisq") seems to show the same significance as marginal hypothesis test

ICtab(m1, m2) #population is a better fit than elevation...
```

```{r}
final_count$predicted_germ_success <- predict(m1, newdata = final_count, type = "response")

final_count$Elev_Level <- cut(final_count$Elev,
                                     breaks = c(-Inf, 786, 1612, 2800),  # Your specified range limits
                                     labels = c("Low", "Medium", "High"))
total_prop <- glm(Germ_prop ~ Elev * Treatment, 
                  family = binomial(link = "logit"), 
                  data = final_count)
total_prop$predicted_total_germ_prop <- predict(total_prop)

predictions <- predict(total_prop, newdata = final_count, type = "response", se.fit = TRUE)
final_count$predicted_total_germ_prop <- plogis(predictions$fit)
final_count$se <- predictions$se.fit
final_count$lower <- plogis(predictions$fit - predictions$se.fit) 
final_count$upper <- plogis(predictions$fit + predictions$se.fit)

ggplot(final_count, aes(x = Elev, y = Germ_prop, color = Treatment)) +
  geom_point() +
  geom_line(aes(y = predicted_total_germ_prop)) +
  ggtitle("Total Germination Proportion After 4 Weeks") +
  labs(x = "Elevation", y = "Total Germination Proportion")
```

```{r}
#ggplot(all_wk_avg, aes(x=Days_to_Germ)) +
 # geom_line(aes(y=Avg_germ_rate2, color=Treatment), size=0.7) +
 # geom_ribbon(aes(ymin = Avg_germ_rate2 - SD_germ_rate2, 
 #                 ymax = Avg_germ_rate2 + SD_germ_rate2, 
 #                 fill = Treatment), 
 #             alpha = 0.1) +
 # theme_bw()+
 # facet_wrap(.~Pop, ncol=4, scales= "free_y")+
 # scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
 # ylab("Average Germination Rate2") +
 # xlab("Days in Inductive Conditions")
```

#### Average germination prop by elevation - this is the average of the total since I filtered to only include the final day proportion

```{r}
all_wk_avg %>%
  filter(Days_to_Germ == 42) %>%
  ggplot(aes(x=Elev, y=Avg_germ_prop, color=Treatment)) +
  geom_point()+
  geom_line()
  #geom_smooth(method = "glm", 
             # method.args = list(family = "binomial"), 
             # se = FALSE, 
             # aes(group = Treatment))
```

```{r}
final_count$predicted <- predict(m1, type = "response", newdata = final_count)

# Create the plot
all_wk_avg %>%
  filter(Days_to_Germ == 42) %>%
  ggplot(aes(x = Elev, y = Avg_germ_prop, color = Treatment)) +
  geom_point() +
  geom_line(data = final_count, aes(x = Elev, y = predicted, color = Treatment), 
            linetype = "solid") 
```

#### Average total germination proportion by elevation - this is the same as what I calculated above

```{r}
all_wk_long %>%
  filter(Days_to_Germ == 42) %>%
  ggplot(aes(x=Elev, y=Avg_total, color=Treatment)) +
  geom_point()+
  geom_line()
```

#### Germ prop ordered by elevation

```{r}
ggplot(all_wk_avg, aes(x=Days_to_Germ)) +
  geom_line(aes(y=Avg_germ_prop, color=Treatment), linewidth=1) +
  geom_ribbon(aes(ymin = Avg_germ_prop - SD_germ_prop, 
                  ymax = Avg_germ_prop + SD_germ_prop, 
                  fill = Treatment), 
              alpha = 0.1) +
  theme_bw()+
  facet_wrap(.~Pop, ncol=4)+
  ylab("Average Germination Proportion") +
  xlab("Days in Inductive Conditions")
```

#### Plot germ rate by pop and treatment ordered by elevation

```{r}
ggplot(all_wk_long, aes(x=Treatment, y=Germ_rate, color=Treatment)) +
  geom_boxplot()+
  theme_bw()+
  facet_wrap(.~Pop, ncol=4, scales="free_y")+
  ylab("Average Germination Rate") 
```

Time to 50% germination
